---
title: "naive bayes penguins"
format: html
editor: visual
---

### Libraries

```{r, message = FALSE}
library(tidyverse)
library(bayesrules)
library(janitor)
library(e1071)
options(scipen = 99)
```

### Story

We’ll start our naive Bayes classification with just a single penguin. Suppose an Antarctic researcher comes across a penguin that weighs less than 4200g with a 195mm-long flipper and 50mm-long bill. Our goal is to help this researcher identify the species of this penguin, Adelie, Chinstrap, or Gentoo.

#### Let's Plot!

Make a visualization that will help the researcher determine the species based on the characteristics described above

```{r}
data(penguins_bayes)
penguins <- penguins_bayes
```

```{r}
ggplot(penguins, aes(x = above_average_weight)) +
  fill = species +
  geom_bar()
```

```{r}
ggplot(penguins, aes(bill_length_mm, flipper_length_mm, color = body_mass_g)) +
  geom_point() +
  geom_smooth(method= "lm") +
  facet_grid(~species)
```

```{r}
#Changes variable name
penguins <- penguins %>%
  mutate(above_average_weight = if_else
        (above_average_weight == 1, "yes", "no"))
```

## Our old bayesian way

P(B\|A)

-   P(Chinstrap \| below average weight)

```{r}
penguins %>%
  tabyl(above_average_weight, species) %>%
  adorn_percentages("row")
```

#### Naive Bayes Classification

```{r}
naive_model_hints <- naiveBayes(species ~ flipper_length_mm + bill_length_mm + above_average_weight, data = penguins)
```

```{r}
our_penguin <- data.frame(bill_length_mm = 50, flipper_length_mm = 195, 
                          above_average_weight= "no")
```

```{r}
predict(naive_model_hints, newdata = our_penguin, type = "raw")
```

#### Test model for accuracy with confusion matrix

```{r}
penguins <- penguins %>%
  mutate(predicted_species = predict(naive_model_hints, newdata = .))
```

```{r}
penguins %>% 
  tabyl(species, predicted_species) %>%
  adorn_percentages("row") %>%
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
naive_model_hints
```

# Lecture 2- Under the hood of Naive Bayes

Species

```{r}
penguins %>%
  tabyl(species)
```

P(adelie) \* P(adelie \| 195mm flipper) \* P(adelie \| 50mm bill length) \* P(under average weight \| adelie) = P (adelie penguin)

### Adelie = 0.016

```{r}
#> Adelie flipper
dnorm(195, mean = 189.9536, sd = 6.539457)
```

```{r}
#> Adelie bill
dnorm(50, mean = 38.79139, sd = 2.663405)
```

```{r}
prob_adelie <- (152/342) * 0.04529578 * 0.999 * 0.83
prob_adelie
```

### Chinstrap = 0.0011

```{r}
#> Chinstrap flipper
dnorm(195, mean = 195.8235, sd = 7.131894)
```

```{r}
#> Chinstrap bill
dnorm(50, mean = 48.83382, sd = 3.339256)
```

```{r}
prob_chinstrap <- (68/342) * 0.05556611 * 0.112 * 0.89
prob_chinstrap
```

### Gentoo = ???

```{r}
#> Gentoo flipper
dnorm(195, mean = 217.1870, sd = 6.484976)
```

```{r}
#> Gentoo bill
dnorm(50, mean = 47.50488, sd = 3.081857)
```

```{r}
prob_gentoo <- (124/342) * 0.0001767016 * 0.09327 * 0.05
prob_gentoo
```

## Sum of probabilities

```{r}
sum_of_probs <- prob_adelie + prob_chinstrap + prob_gentoo
sum_of_probs
```

### Probability of chinstrap penguin

```{r}
prob_chinstrap / sum_of_probs
```
